Efficient Community Formation \\ for Web Services

### 2013-12-15 papers/coalition-journal-R1/Coalition_Journal_JB_Sep1.tex
ehsan

\usepackage{amsfonts}
\usepackage{amssymb,amsmath}

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Efficient Coalition Formation \\ for Web Services}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
        John~Doe,~\IEEEmembership{Fellow,~OSA,}
        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://www.michaelshell.org/contact.html
\IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Web services are loosely-coupled business applications willing to
cooperate in distributed settings within different groups called
communities. Communities aim to provide better visibility,
efficiency, market share and total payoff. There are a number of
proposed mechanisms and models on aggregating web services and
making them cooperate within their communities. However, forming
optimal and stable communities as coalitions to maximize
individual and group efficiency and income has not been addressed
yet. In this paper, we propose an efficient coalition formation
mechanism using cooperative game-theoretic techniques. We propose
a mechanism for community membership requests and selections of
web services under two scenarios where: (1) there is interaction between one community and many web services
and (2) web services can join multiple established communities. The ultimate objective is to
develop a mechanism for web services to form stable groups
allowing them to maximize their efficiency and generate
near-optimal (welfare-maximizing) communities. The theoretical and
extensive simulation results show that our algorithms provide web services
and community owners, in real-world like environments, with applicable and near-optimal decision
making mechanisms.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

@@

\IEEEPARstart{O}{ver} the past years, online services have become part of many
scalable business applications. The increasing reliance on
web-based applications has significantly influenced the way web
services are engineered. Web services provide a set of stateless
software functions accessible at a network address over the web.
The recent developments are shifting web services from passive and
individual components to autonomous and group-based components
where interaction, composition, and cooperation are the key
challenges \cite{ICWS2011-1,SCC2011-1}. The main objective is to
achieve a seamless integration of business processes, applications
and web services. Delivering high quality services considering the
dynamic and unpredictable nature of the Internet is still a very
critical and challenging issue.

The need for highly available and responsive services has called
for grouping and collaborative mechanisms of loosely-coupled web
services, particularly in business settings. The idea of grouping
web services within communities and the way those communities are
engineered so that web services can better collaborate have been
proposed and investigated in
\cite{DBLP:journals/ijebr/MaamarSTBB09,DBLP:journals/internet/BenatallahSD03,Rosario:2008:PQS:1512146.1512290}.
Communities are virtual groups of web services having similar
functionalities \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038,Medjahed05adynamic,10.1109/ARES.2008.7}, but probably different non-functional quality
attributes, which form the QoS parameters. When communities are
used, users send their requests to the masters of those
communities, which are responsible of managing the communities,
forwarding the requests to the suitable member web services and
checking the credentials of those members. Communities aim to
provide higher service availability and performance than what
individual web services can provide. The high availability of
services and the community resilience to failure are guaranteed
since web services can cooperate and replace each other within the
same community and since there is no single point of failure in
the communities architecture.

@@

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed mechanisms to help
users and services to maximize their gain. A two-player
non-cooperative game between web services and community master was
introduced in \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this
game-theoretic model, the strategies available to a web service
when facing a new community are requesting to join the community,
accepting the master's invitation to join the community, or
refusing the invitation to join. The set of strategies for
communities are inviting the web service or refusing the web
service's join request. Based on their capacity, market share and
reputation, the two players have different set of utilities over
the strategy profiles of the game. The main limits of this game
model are: 1) its consideration of only three quality parameters,
while the other factors are simply ignored; and 2) the
non-consideration of the web services already residing within the
community. The game is only between the community master and the
new web service, and the inputs from all the other members are
simply ignored. The consideration of those inputs is a significant
issue as existing web services can lose utility or payoff because
of the new member, which can results in an unhealthy and unstable
group. The problem comes from the fact that the existing members
should collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

@@

In \cite{10.1109/TSC.2012.12}, a cooperative scheme among autonomous
web services based on coalitional game theory has been introduced. They have proposed an algorithm to
reach individually stable coalition partition for web services in order to
maximize their efficiency. The communities choose new web services on the promise
that it would benefit the community without decreasing any other web service's
income. In their model, the worth of community is evaluated with high emphasis on
availability metric and considering price and cost values only. The community structure is based on a coordination chain,
where a web service is assigned as a \emph{primary} web service and the community task destribution
method, will initially invoke the primary web service and only if the primary web service is unavailable
will invoke the next backup web services as they are ordered in the coordination chain. However in cooperative models, it is preferred to
have a real and active cooperative activity engaging all agents to perform the tasks more efficiently. Especially nowadays
with recent advancement in cloud and hardware infrastructures availability is becoming less of an issue. So the backup web services
in their model have a very low chance of getting jobs, especially the ones further in chain, which is huge waste of web services
capabilities.

\textbf{Contributions.} In this paper, we use game theory to
propose a cooperative game model for the aggregation of web
services within communities. The solution concepts of our
cooperative game seeks to find efficient ways of forming
coalitions (teams) of web services so that they can maximize their
gain and payoff, and distribute the gain in a fair way among all
the web services. Achieving Fairness when the gain is distributed
among the community members is the main factor to keep the
coalition stable as no web service will expect to gain better by
deviating for the community. In other words, the coalition is made
efficient if all the members are satisfied. We first propose a
representation function for communities of web services based on
their QoS attributes. By using this function, we can evaluate the
$worth$ of each community of web services. When facing new
membership requests, a typical community master checks whether the
new coalition having the old and new set of web services will keep
the community stable or not. The community master will reject the
membership requests if it finds out that the new coalition would
be unstable, preventing $any$ subset of web services from gaining
significantly more by deviating from the community and joining
other communities or forming new ones. The computation of
solutions for cooperative game theory problems is combinatorial in
nature and proven to be NP-complete \cite{Algorithmic}, making
this computation impractical in real world applications. However,
using the concepts of coalition stability, we propose
approximation algorithms running in polynomial time providing web
services and community masters with applicable and near-optimal
decision making mechanisms.

The rest of paper is organized as follows. Section
\ref{s:preliminaries} describes the architecture, considered
parameters and some basic cooperative game theory concepts.
Section \ref{s:model} provides the problem statement and some numerical examples.
Section \ref{s:game_solution} presents our solution model in two
different scenarios. Section \ref{s:resutls} describes our
experiments and results. Finally, Section \ref{s:conclusion},
concludes the paper and identifies future work.

@@

\emph{1) Web services} are rational entities\footnote{The term
rational is used here in the sense that web services are utility
maximizers} providing services to end users. They aim to maximize
their individual income by receiving enough requests from end
users. In order to increase their revenue, web services seek for
more tasks if they have the capacity and throughput to do so. Web
services can join communities to have better efficiency by
collaborating with others, to have access to higher market share,
and to have opportunity of receiving a bigger task pool from end
users. Throughout this paper, in
our equations, we refer to web services as $ws$ and to the set of
web services hosted by a given community as $C$. To simplify the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

@@

\emph{3) Users} generate requests and try to find the best
available services. User satisfaction is abstracted as function of
quantity and quality of tasks accomplished by a given service. 
Higher user satisfaction leads to higher trust of the community by users hence directing more requests towards that service provider.

@@

\begin{table}[!t]
\centering
\caption{List of web service QoS parameters.}
\begin{tabular}{|c|c||c|c|}
\hline
\textbf{Parameter} & \textbf{Definition} \\
\hline\hline
$Availability$ & Probability of being available during \\
&a time frame \\
$Reliability$ & Probability of successfully handling \\
&requests during a timeframe\\
$Successability$ & Rate of successfully handled requests \\
$Throughput$ & Average rate of handling requests \\
$Latency$ & The average latency of services\\
$Capacity$ & Amount of resources available\\
$Cost$ & Mean service fee \\
$Regulatory$ & Compliance with standards, law and rules\\
$Security$ & Quality of confidentiality \\
&and non-repudiation\\
\hline
\end{tabular}
\label{qosws}
\end{table}

@@

Figure \ref{fig_community} represents the architecture of web service communities. The communities are essentially an abstract model of web services. They aggregate web services and communicate with other entities such as UDDI registries and users, using identical protocols as web services. Web services join communities to increase their utility by having a larger market share and task pool. Community coordinators or master web services are responsible for community development, managing membership requests from web services and distributing user tasks among the community members. Community coordinators try to attract quality web services and keep the community as stable and productive as possible to gain better reputation and user satisfaction which results in having a higher market share for the community. The way the web services reside inside communities and how communities of web services are engineered is described comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\begin{figure*}[!t]
\centerline{\includegraphics[width=6.25in]{community.eps}}
\caption{Architecture of Web Service communities}
\label{fig_community}
\end{figure*}

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players which can form a group called a $coalition$. A
\emph{coalitional game} is a pair $G = (N, v)$, where $v$ is called
a \emph{characteristic function} $v: 2^N \to \mathbb{R}$, mapping the set of players of the
coalition to a real number $v(N)$, the worth of $N$. This number
usually represents the output or payoff or again the performance
of these players working together as coalition.  If a coalition
$S$ is formed, then it can divide its worth, $v(N)$ in any
possible way among its members. The payoff vector $x \in
\mathbb{R}^N$ is the amount of payoff being distributed among the
members of the coalition $N$. The payoff vector satisfies two
conditions:

@@

When dealing with cooperative games, two issues need to be
addressed:\\ 1. Which coalitions among all possible coalitions to form? \\
2. How to reward each member when a task is completed?\\
%
The following definitions help address these two issues.

@@

{\bf Definition 1 (Shapley value)} Given a cooperative game $(N,
v)$, the \emph{Shapley value} of player $i$ is given
by\cite{shapley_value}:
\begin{equation}\label{eq:shapley}
\phi_i(N,v) = \sum_{S \subseteq N \backslash \left\{i\right\} }
\frac{|S|! (|N|-|S|-1)!}{|N|!} (v(S \cup \left\{i\right\}) - v(S))
\end{equation}

\emph{Shapley value} is a unique and fair solution concept for
payoff distribution among the members of the coalition. It
basically rewards members with the amount of average marginal contribution
they add to the coalition. It checks the contribution of member $i$ by adding the agent, to all possible subsets 
of coalitions $S$, where ${S \subseteq N\left\backslash\left\{i\right\}$. If he is added to the set $S$, his 
contribution to the coalition is $v(S \cup \left\{i\right\}) - v(S)$. Average marginal contribution of agent $i$'s 
is calculated by averaging this value over all possible subsets of $N$, in $Shapley value$ equation (\ref{eq:shapley}).

@@

Coalition structure formation is the problem of finding the best
partition of web services into teams. In these settings, the
performance of an individual service is less important than the
\emph{social welfare} of the whole system, which is the sum of the
values of all teams. Having the game $(N,v)$, a coalition
structure $(CS)$ is \emph{socially optimal} if $CS$ belongs to set
$\operatorname*{arg\,max}_{CS} v(CS)$ where $v(CS)$ is the sum of
the values of all coalitions inside $CS$. $v(CS) = \sum_{C \in
CS}v(C)$.
%The outcome of a characteristic function game in coalition structure settings, consists of two parts; first a disjoint partition of players (agents) into coalitions, called a \emph{coalition structure} (CS) and second a \emph{payoff vector} as mentioned in cooperative game solution concepts, which distributes the value of each coalition among its members.

>>>>

\usepackage{amsfonts}
\usepackage{amssymb,amsmath}
 \usepackage{comment}

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Efficient Community Formation \\ for Web Services}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Ehsan~Khosrowshahi~Asl,~\IEEEmembership{Member,~IEEE,}
        Jamal~Bentahar,~\IEEEmembership{Member,~IEEE,}
        Hadi~Otrok,~\IEEEmembership{Member,~IEEE,}
        and~Rabeb~Mizouni,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell is with
the Department of Electrical and Computer Engineering, Georgia
Institute of Technology, Atlanta,
GA, 30332.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://www.michaelshell.org/contact.html
\IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Web services are loosely-coupled business applications willing to
cooperate in distributed settings within different groups called
communities. Communities aim to provide web services with better
visibility, efficiency, market share and total payoff. A number of
mechanisms and models have been recently proposed to aggregate web
services and make them cooperate within their communities.
However, forming optimal and stable communities as coalitions to
maximize individual and group efficiency and income is yet to be
addressed. In this paper, we propose an efficient community
formation mechanism using cooperative game-theoretic techniques,
particularly Shapley value, core, $\epsilon$-core and convex
games. We propose a mechanism for community membership requests
and selections of web services under two scenarios: (1) there is
only one community and many web services aim to join it; and (2)
web services can join different established communities. The
ultimate objective is to develop a mechanism for web services to
form stable groups allowing them to maximize their efficiency and
generate near-optimal (welfare-maximizing) communities where a
taxation-based solution is being proposed. The theoretical and
extensive simulation results show that our algorithms provide web
services and community owners, in real-world-like environments,
with applicable and near-optimal decision making mechanisms.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

@@

\IEEEPARstart{O}{ver} the past years, online services have become
part of many scalable business applications. The increasing
reliance on web-based applications has significantly influenced
the way web services are engineered. Web services provide a set of
stateless software functions accessible at a network address over
the web. The recent developments are shifting web services from
passive and individual components to autonomous and group-based
components where interaction, composition, and cooperation are the
key challenges \cite{ICWS2011-1,SCC2011-1,Baldoni2010,Younas2013}.
The main objective is to achieve a seamless integration of
business processes, applications and web services. Delivering high
quality services considering the dynamic and unpredictable nature
of the Internet is still a very critical and challenging issue.

The need for highly available and responsive services has called
for grouping and collaborative mechanisms of loosely-coupled web
services, particularly in business settings
\cite{YuB13,Younas2012,LiuBWZ13}. The idea of grouping web
services within communities and the way those communities are
engineered so that web services can better collaborate have been
proposed and investigated in
\cite{DBLP:journals/ijebr/MaamarSTBB09,DBLP:journals/internet/BenatallahSD03,Rosario:2008:PQS:1512146.1512290}.
Communities are virtual groups of web services having similar or
complementary functionalities \cite{Zeng:2003:QDW:775152.775211,
Paik:2005:TSS:2229263.2230038,Medjahed05adynamic,10.1109/ARES.2008.7},
but probably different non-functional quality attributes, which
form the QoS parameters. When communities are used, users send
their requests to the masters of those communities, which are
responsible of managing the communities, forwarding the requests
to the suitable member web services and checking the credentials
of those members. Communities aim to provide higher service
availability and performance than what individual web services can
do. The high availability of services and the community resilience
to failure are guaranteed since web services can cooperate and
replace each other within the same community and since there is no
single point of failure in the communities architecture.

@@

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community
master was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this game-theoretic
model, the strategies available to a web service when facing a new
community are requesting to join the community, accepting the
master's invitation to join the community, or refusing the
invitation to join. The set of strategies for communities are
inviting the web service or refusing the web service's join
request. Based on their capacity, market share and reputation, the
two players have different sets of utilities over the strategy
profiles of the game. The main limits of this game model are: 1)
its consideration of only three quality parameters, while the
other factors are simply ignored; and 2) the non-consideration of
the web services already residing within the community. The game
is only between the community master and the new web service, and
the inputs from all the other members and their influence on the
master's decision are simply ignored. The consideration of those
inputs and this influence factor is a significant issue as
existing web services can lose utility or payoff because of the
new member, which can results in an unhealthy and unstable group.
The problem comes from the fact that the existing members should
collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

@@

In \cite{10.1109/TSC.2012.12}, a cooperative scheme among
autonomous web services based on coalitional game theory has been
introduced. The authors have proposed an interesting algorithm to
reach individually stable coalition partition for web services in
order to maximize their efficiency. The communities choose new web
services on the promise that it would benefit the community
without decreasing any other web service's income. In the proposed
model, the worth of community is evaluated with high emphasis on
the availability metric and considering price and cost values
only. The community structure is based on a coordination chain,
where a web service is considered as a \emph{primary} web service
and the community task-distribution method initially invokes the
primary web service and only if the primary web service is
unavailable, the method invokes the next backup web services as
they are ordered in the coordination chain. We believe that this
coordination chain limits the cooperation power as it introduces a
sort of hierarchy. However, in pure and open cooperative models,
such as the one we propose in this paper, active cooperation
activities engaging simultaneously many agents so that they can
perform the tasks more efficiently are being used. Moreover, if
the availability is high, which is the case nowadays with the
recent advancements in cloud and hardware infrastructures, the
backup web services will end-up having a very low chance of
getting jobs, especially the ones further in the chain. This will
results in a considerable waste of web services capabilities.

\textbf{Contributions.} In this paper, we use game theory to
propose a cooperative game model for the aggregation of web
services within communities. The solution concepts of our
cooperative game seeks to find efficient ways of forming
coalitions (teams) of web services so that they can maximize their
gain and payoff, and distribute the gain in a fair way among all
the web services. Achieving fairness when the gain is distributed
among the community members is the main factor to keep the
coalition stable as no web service will expect to gain better by
deviating from the community. In other words, the coalition is
made efficient if all the members are satisfied. We analyze the
game model using two scenarios. The first one simply considers one
community and many web services interested to join it. The second
scenario, which generalizes the first one, includes the
possibility of joining different already established communities.
We first propose a representation function for communities of web
services based on their QoS attributes. By using this function, we
can evaluate the $worth$ of each community of web services. When
facing new membership requests, a typical community master checks
whether the new coalition having the old and new set of web
services will keep the community stable or not. The community
master will reject the membership requests if it finds out that
the new coalition would be unstable, preventing $any$ subset of
web services from gaining significantly more by deviating from the
community and joining other communities or forming new ones. The
computation of solutions for cooperative game theory problems is
combinatorial in nature and proven to be NP-complete
\cite{Algorithmic}, making this computation impractical in real
world applications. However, using the concepts of coalition
stability, we propose approximation algorithms running in
polynomial time providing web services and community masters with
applicable and near-optimal decision making mechanisms.

The rest of the paper is organized as follows. Section
\ref{s:preliminaries} describes the architecture, considered
parameters and some basic cooperative game theory concepts.
Section \ref{s:model} provides the problem statement and some
numerical examples for illustration. Section \ref{s:game_solution}
presents our solution model in two different scenarios. Section
\ref{s:resutls} describes our experiments and results. Finally,
Section \ref{s:conclusion} concludes the paper and identifies
future work.

@@

\emph{1) Web services} are rational entities\footnote{The term
rational is used here in the sense that web services are utility
maximizers} providing services to end users. They aim to maximize
their individual income by receiving enough requests from end
users. In order to increase their revenue, web services seek for
more tasks if they have the capacity and throughput to do so. Web
services can join communities to have better efficiency by
collaborating with others, to have access to higher market share,
and to have opportunity of receiving a bigger task pool from end
users. Throughout this paper, in our equations, we refer to web
services as $ws$ and to the set of web services hosted by a given
community as $C$. To simplify the notation, sometimes we simply
write $ws$ instead of $ws \in C$ to go through the elements $ws$
of the set $C$, when $C$ can be understood from the context.

@@

\emph{3) Users} generate requests and try to find the best
available services. User satisfaction is abstracted as function of
quantity and quality of tasks accomplished by a given service.
Higher user satisfaction leads to higher trust an better
reputation of the community as a whole, which helps direct more
requests towards that community.

@@

\begin{table}[!t]
\centering
\caption{List of web service QoS parameters.}
\begin{tabular}{|c|c||c|c|}
\hline
\textbf{Parameter} & \textbf{Definition} \\
\hline\hline
$Availability$ & Probability of being available during \\
&a time frame \\
$Reliability$ & Probability of successfully handling \\
&requests during a timeframe\\
$Successability$ & Rate of successfully handled requests \\
$Throughput$ & Average rate of handling requests \\
$Latency$ & Average response round-trip time\\
$Capacity$ & Amount of resources available\\
$Cost$ & Mean service fee \\
$Regulatory$ & Compliance with standards, law and rules\\
$Security$ & Quality of confidentiality \\
&and non-repudiation\\
\hline
\end{tabular}
\label{qosws}
\end{table}

@@

Figure \ref{fig_community} represents the architecture of web
service communities. The community is essentially a virtual group
of web services having similar and complementary functionalities.
Communities aggregate web services and communicate with other
entities such as UDDI registries and users, using identical
protocols as single web services. Web services join communities to
increase their utility by having a larger market share and task
pool. Community coordinators or master web services are
responsible for community development, managing membership
requests from web services and distributing user tasks among the
community members. Community coordinators try to attract quality
web services for joining and keep the community as stable and
productive as possible to gain better reputation and user
satisfaction, which results in having a higher market share for
the community. The way the web services reside inside communities
and how communities of web services are engineered is described
comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\begin{figure*}[!t]
\centerline{\includegraphics[width=6.25in]{community.eps}}
\caption{Architecture of web service communities}
\label{fig_community}
\end{figure*}

\subsection{Cooperative Game Concepts}
The theory of cooperative games is a branch of game theory that
studies strategies of self-interested entities or agents in a
setting where those agents can increase their payoff by binding
agreements and cooperating in groups. We let $N$ be a set of
players. Any subset $S$ of $N$ can form a group called
$coalition$. A \emph{coalitional game} is a pair $G = (N, v)$,
where $v$ is called a \emph{characteristic function} $v: 2^N \to
\mathbb{R}$, mapping the set of players of the coalition to a real
number $v(S)$, the worth of $S$. This number usually represents
the output, payoff, or the performance of these players working
together as coalition.  If a coalition $S$ is formed, then it can
divide its worth, $v(S)$ in any possible way among its members.
The payoff vector $x \in \mathbb{R}^S$ is the amount of payoff
being distributed among the members of the coalition $S$. The
payoff vector satisfies two conditions:

@@

When dealing with cooperative games, two issues need to be
addressed  \cite{
DBLP:conf/ijcai/GrecoMPS11,Sandholm1999209,DBLP:conf/ijcai/RahwanMJ11}:\\ 1. Which coalitions among all possible coalitions to form? \\
2. How to reward each member when a task is completed?\\
%
The following definitions help address these two issues
\cite{shapley_value}, \cite{S1971cores}.

@@

{\bf Definition 1 (Shapley value)} Given a cooperative game $(N,
v)$, the \emph{Shapley value} of player $i$ is given by:
\begin{equation}\label{eq:shapley}
\phi_i(N,v) = \sum_{S \subseteq N \backslash \left\{i\right\} }
\frac{|S|! (|N|-|S|-1)!}{|N|!} (v(S \cup \left\{i\right\}) - v(S))
\end{equation}

\emph{Shapley value} is a unique and fair solution concept for
payoff distribution among the members of the coalition. It
basically rewards members with the amount of average marginal
contribution they add to the coalition. It checks the contribution
of member $i$ by adding the agent to all possible subsets of
coalitions $S$, where $S \subseteq N-\{i\}$. If he is added to the
set $S$, his contribution to the coalition is $v(S \cup
\left\{i\right\}) - v(S)$. Average marginal contribution of agent
$i$'s is calculated by averaging this value over all possible
subsets of $N$, in the Shapley value equation (Equation
\ref{eq:shapley}).

@@

Coalition structure formation is the problem of finding the best
partition of web services into teams (i.e., communities). In these
settings, the performance of an individual service is less
important than the \emph{social welfare} of the whole system,
which is the sum of the values of all communities. Having the game
$(N,v)$, a coalition structure $(CS)$ is \emph{socially optimal}
if $CS$ belongs to set $\operatorname*{arg\,max}_{CS} v(CS)$ where
$v(CS)$ is the sum of the values of all coalitions inside $CS$.
$v(CS) = \sum_{C \in CS}v(C)$.
%The outcome of a characteristic function game in coalition structure settings, consists of two parts; first a disjoint partition of players (agents) into coalitions, called a \emph{coalition structure} (CS) and second a \emph{payoff vector} as mentioned in cooperative game solution concepts, which distributes the value of each coalition among its members.

In this section, we present  web services and community
coordinator's interactions, the task distribution process, and
revenue models in web service communities.

\subsection{Task Distribution}

As mentioned in Section \ref{webservice-communities}, communities
are robust service providers with well established market share
and reputation. By maintaining their reputation and performance,
they attract  end users which choose them as service providers to
perform their tasks. The community master is characterized by a
request rate (also called input rate) $(R_C)$ from users. Each web
service comes with a given QoS ($QoS_{ws}$) from which the
throughput $Th_{ws}$ is excluded. Throughput is the average rate
of tasks a web service can perform per time unit. Its exclusion
from $QoS_{ws}$ allows us to build our analysis on the particular
value of $Th_{ws}$. Thus, web services perform tasks with an
average output quality of $QoS_{ws}$ and a throughput rate of
$Th_{ws}$.

The community master uses a slightly modified \emph{weighted fair
queuing} method to distribute tasks among its members. The goal is
to allocate incoming tasks to web services with a rate matching
the throughput value of $Th_{ws}$. In the \emph{weighted fair
queuing} method, \emph{all} the input flow is multiplexed along
different paths, however in our case if the request rate $(R_C)$
of the community is more than the summation of throughput values
of the web services in the community, some of the input tasks will
be queued and served with delay. Thus, the amount of tasks
performed by the community is $\sum_{ws}{Th_{ws}}$ when
$\sum_{ws}{Th_{ws}} \leq R_{C}$. However, when the input rate
$(R_C)$ of the community is less than the summation of throughput
values of the web services in the community,
%the community has more web services having more total throughput value than community's request rate
the \emph{weighted fair queuing} algorithm assigns a weighted task
rate of $R_C \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each
web service ($ws$), so that the total rate of tasks being
performed is $R_C$, the community's request rate.

While distributing tasks, the community master can verify the
performance, throughput and quality of service of tasks being
performed by web services. The master can recognize if those web
services are capable of doing the amount of tasks they advertised.
If for any reason there is a decline in the quality metric or
throughput, the master will announce the new parameters, which
will be considered as benchmark for future performance
calculations, and penalize the suspicious web services. In this
way, players will have incentive to truthfully reveal their real
capabilities in order to profit best from the community and to
avoid being penalized. In addition, the system should be dynamic
enough to detect and react to web services quality metrics
variation as over time web service metrics may degrade or improve,
a change that the community should adjust to.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

The communities and web services earn revenue by performing tasks.
The total gain is function of quality ($QoS_{ws}$) and throughput
($Th_{ws}$) of tasks being performed. As mentioned in Section
\ref{ws_parameters}, $QoS_{ws}$ is obtained by a certain
aggregation function of the parameters considered in Table
\ref{qosws}. We have adopted a linear equal weight average over
all QoS parameters listed in Table \ref{qosws} excluding the
$Throughput$ and $Cost$ parameters. A community has the option to
weigh specific QoS parameters depending on the expectations of
their clients.

The maximum potential output of a community $C$ $(PO(C))$  is an
aggregation of the number of performed tasks $T_{ws}$ times the
quality of service $QoS_{ws}$ for each web service member of the
community:

If the summation of throughput values ($Th_{ws}$) of community
members exceeds the input task rate of the community ($R_C$), the
community cannot perform at its maximum potential. This situation
denotes the case when the community has more web services than it
needs to perform the input task load. Thus, the actual output has
to be normalized to the amount of tasks being performed.

\subsection{Illustrative Examples}

In this section, we analyze three numerical examples and discuss
the motivation of web services and community interactions, the
strategies they can adopt, and the revenue they can earn by
adopting these different strategies.

%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Example 1} \label{example_1} \centering
\begin{tabular}{c c c c}
\hline
$WS$ & $QoS_{ws}$ & $Th_{ws}$ & $Th_{ws} \times QoS_{ws}$\\
\hline
1 & 0.8 & 4 & 3.2\\
2 & 0.8 & 5 & 4.0\\
3 & 0.8 & 3 & 2.4\\
\hline\\
\end{tabular}
%\end{table}
%
%\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
% \caption{Three web services}
\label{example_1_2} \centering
\begin{tabular}{c c || c c}
\hline
Community & Worth & Community & Worth\\
\hline
$\left\{1\right\}$ & 3.2 & $\left\{1,2\right\}$ & 7.2\\
$\left\{2\right\}$ & 4.0 & $\left\{1,3\right\}$ & 5.6\\
$\left\{3\right\}$ & 2.4 & $\left\{2,3\right\}$ & 6.4\\
$\left\{1,2,3\right\}$ & 8.0\\
\hline
Community $R_C$: 10\\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
